---
title: "Unjournal Evaluations Data - January 2026"
description: "Static snapshot of evaluation data while the interactive dashboard is being updated"
date: "2026-01-19"
categories: [data, evaluations]
---

```{r}
#| label: setup
#| include: false

library(dplyr)
library(ggplot2)
library(forcats)
library(stringr)
library(knitr)
library(here)

research <- readr::read_csv(here("data/research.csv"), show_col_types = FALSE)
ratings <- readr::read_csv(here("data/rsx_evalr_rating.csv"), show_col_types = FALSE)

res_ratings <- inner_join(ratings, research,
                          by = join_by(research == label_paper_title))

my_theme <- theme_minimal() +
      theme(
        legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.y = element_text(size = 12),
        axis.text.x = element_text(size = 10),
        plot.title = element_text(size = 14)
      )
theme_set(my_theme)

criteria_titles <- c(
  "overall"         = "Overall assessment",
  "adv_knowledge"   = "Advancing our knowledge and practice",
  "methods"         = "Methods",
  "logic_comms"     = "Logic and communication",
  "real_world"      = "Real world relevance",
  "gp_relevance"    = "Relevance to global priorities",
  "claims"          = "Claims and evidence",
  "open_sci"        = "Open, collaborative, replicable science",
  "merits_journal"  = "Deserved journal tier",
  "journal_predict" = "Predicted journal tier"
)
```

::: {.callout-note}
This page provides a static snapshot of Unjournal evaluation data while we update the interactive dashboard. Visit the [main dashboard](https://unjournal.shinyapps.io/uj-dashboard) for the interactive version.
:::

## Summary Statistics

```{r}
#| label: summary
#| echo: false

evaled <- research |>
  filter(status == "50_published evaluations (on PubPub, by Unjournal)")

n_papers <- nrow(evaled)
n_evals <- n_distinct(select(res_ratings, research, evaluator))
n_areas <- n_distinct(evaled$main_cause_cat_abbrev, na.rm = TRUE)
```

- **`r n_evals`** evaluations of **`r n_papers`** papers across **`r n_areas`** research areas

## Research Areas

```{r}
#| label: areas-plot
#| echo: false
#| fig-width: 8
#| fig-height: 5

evaled |>
  filter(!is.na(main_cause_cat_abbrev)) |>
  mutate(
    area = str_to_title(main_cause_cat_abbrev),
    area = fct_infreq(area),
    area = fct_rev(area)
  ) |>
  ggplot(aes(y = area, fill = area)) +
    geom_bar() +
    labs(
      title = "Research areas of evaluated papers",
      x = "Number of papers",
      y = NULL
    )
```

### Research Areas Table

```{r}
#| label: areas-table
#| echo: false

evaled |>
  filter(!is.na(main_cause_cat_abbrev)) |>
  count(Area = str_to_title(main_cause_cat_abbrev), name = "Papers") |>
  arrange(desc(Papers)) |>
  kable()
```

## Overall Ratings by Research Area

```{r}
#| label: ratings-plot
#| echo: false
#| fig-width: 9
#| fig-height: 6

plot_data <- res_ratings |>
  filter(criteria == "overall") |>
  filter(!is.na(main_cause_cat_abbrev)) |>
  mutate(
    area = str_to_title(main_cause_cat_abbrev),
    area = fct_infreq(area),
    area = fct_rev(area)
  )

means <- plot_data |>
  summarize(mean_rating = mean(middle_rating, na.rm = TRUE), .by = area)

ggplot(plot_data, aes(y = area, x = middle_rating)) +
  geom_jitter(aes(color = area), height = 0.2, width = 0, alpha = 0.7, size = 3) +
  geom_point(data = means, aes(x = mean_rating, y = area),
             shape = 4, size = 4, stroke = 2, color = "black", inherit.aes = FALSE) +
  scale_x_continuous(breaks = seq(0, 100, 20), limits = c(0, 100)) +
  labs(
    title = "Overall ratings of papers by research area",
    subtitle = "Points = individual papers, X = area average",
    x = "Rating (0-100 percentile)",
    y = NULL
  )
```

### Overall Ratings Table

```{r}
#| label: ratings-table
#| echo: false

res_ratings |>
  filter(criteria == "overall") |>
  select(Paper = research, Area = main_cause_cat_abbrev,
         Rating = middle_rating, `Lower CI` = lower_CI, `Upper CI` = upper_CI) |>
  mutate(Area = str_to_title(Area)) |>
  arrange(desc(Rating)) |>
  kable(digits = 1)
```

## Ratings by Evaluation Criteria

```{r}
#| label: criteria-plot
#| echo: false
#| fig-width: 9
#| fig-height: 7

criteria_data <- res_ratings |>
  filter(criteria %in% names(criteria_titles)) |>
  filter(!criteria %in% c("merits_journal", "journal_predict")) |>
  mutate(
    criteria_label = criteria_titles[criteria],
    criteria_label = fct_relevel(criteria_label, "Overall assessment"),
    criteria_label = fct_rev(criteria_label)
  )

means <- criteria_data |>
  group_by(criteria_label) |>
  summarize(mean_rating = mean(middle_rating, na.rm = TRUE))

ggplot(criteria_data, aes(y = criteria_label, x = middle_rating, color = criteria_label)) +
  geom_jitter(height = 0.2, width = 0, alpha = 0.6, size = 2) +
  geom_point(data = means, aes(x = mean_rating, y = criteria_label),
             shape = 4, size = 3, stroke = 2, color = "black", inherit.aes = FALSE) +
  scale_x_continuous(breaks = seq(0, 100, 20), limits = c(0, 100)) +
  labs(
    title = "Ratings by evaluation criteria (0-100 scale)",
    subtitle = "Points = individual ratings, X = criterion average",
    x = "Rating",
    y = NULL
  )
```

### Criteria Averages Table

```{r}
#| label: criteria-table
#| echo: false

res_ratings |>
  filter(criteria %in% names(criteria_titles)) |>
  mutate(Criteria = criteria_titles[criteria]) |>
  group_by(Criteria) |>
  summarize(
    `Mean` = mean(middle_rating, na.rm = TRUE),
    `Median` = median(middle_rating, na.rm = TRUE),
    `Min` = min(middle_rating, na.rm = TRUE),
    `Max` = max(middle_rating, na.rm = TRUE),
    `N` = n()
  ) |>
  arrange(desc(Mean)) |>
  kable(digits = 1)
```

## Paper Sources

```{r}
#| label: sources-plot
#| echo: false
#| fig-width: 8
#| fig-height: 5

evaled |>
  mutate(
    Source = case_match(source_main,
      "internal-NBER"                ~ "Internal (NBER)",
      "internal-from-syllabus-agenda-policy-database" ~ "Internal",
      "suggested - internally"       ~ "Internal",
      "submitted (by author(s))"     ~ "Author",
      "suggested - externally"       ~ "Third party",
      "suggested - externally - NGO" ~ "Third party (NGO)",
      .default = "Unknown"
    ),
    Source = fct_infreq(Source),
    Source = fct_rev(Source)
  ) |>
  ggplot(aes(y = Source, fill = Source)) +
    geom_bar() +
    labs(
      title = "Sources of evaluated papers",
      x = "Number of papers",
      y = NULL
    )
```

### Sources Table

```{r}
#| label: sources-table
#| echo: false

evaled |>
  mutate(
    Source = case_match(source_main,
      "internal-NBER"                ~ "Internal (NBER)",
      "internal-from-syllabus-agenda-policy-database" ~ "Internal",
      "suggested - internally"       ~ "Internal",
      "submitted (by author(s))"     ~ "Author",
      "suggested - externally"       ~ "Third party",
      "suggested - externally - NGO" ~ "Third party (NGO)",
      .default = "Unknown"
    )
  ) |>
  count(Source, name = "Papers") |>
  arrange(desc(Papers)) |>
  kable()
```

## All Evaluated Papers

```{r}
#| label: all-papers
#| echo: false

evaled |>
  select(Paper = label_paper_title, Area = main_cause_cat_abbrev,
         Status = publication_status, Source = source_main) |>
  mutate(Area = str_to_title(Area)) |>
  arrange(Paper) |>
  kable()
```

---

*Data last updated: `r format(Sys.time(), "%Y-%m-%d %H:%M")`*
