---
title: "The Unjournal data dashboard"
server: shiny
format:
  dashboard:
    theme: simplex
    expandable: false
---

```{r}
#| label: setup
#| context: setup
#| include: false

library(conflicted)
library(cowplot)
library(dplyr)
conflicts_prefer(dplyr::filter)
library(forcats)
library(gdtools)
library(gfonts)
library(ggplot2)
library(here)
library(lubridate)
library(shiny)
library(stringr)
library(tidyr)

register_gfont("Open Sans")
addGFontHtmlDependency(family = "Open Sans")
# names are display periods; values are relevant starting dates for data
periods <- c(
  "Ever"      = as.Date("1900-01-01"),
  "12 months" = today() - period(12, "months"),
  "6 months"  = today() - period(6, "months"),
  "3 months"  = today() - period(3, "months"),
  "1 month"   = today() - period(1, "months")
)


my_theme <- theme_minimal() +
      theme(
        legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.y = element_text(size = 16),
        axis.text.x = element_text(size = 14),
        plot.title = element_text(size = 18)
      )
ggplot2::theme_set(my_theme)

criteria_titles <- c(
        "overall"         = "Overall assessment",
        "adv_knowledge"   = "Advancing our knowledge and practice",
        "methods"         = "Methods",
        "logic_comms"     = "Logic and communication",
        "real_world"      = "Real world relevance",
        "gp_relevance"    = "Relevance to global priorities/usefulness to practitioners",
        "claims"          = "Claims, strength, and characterization of Evidence",
        "open_sci"        = "Open, collaborative, replicable science",
        "merits_journal"  = "Deserved journal tier",
        "journal_predict" = "Predicted journal tier"
      )
```


```{r}
#| label: data-import
#| context: data
#| include: false

research <- readr::read_csv(here("data/research.csv"))
ratings <- readr::read_csv(here("data/rsx_evalr_rating.csv"))

res_ratings <- inner_join(ratings, research, 
                          by = join_by(research == label_paper_title))

research_areas <- unique(res_ratings$main_cause_cat_abbrev)
research_areas <- sort(research_areas)
research_areas <- c("All", research_areas)
names(research_areas) <- research_areas

```


## {.toolbar}

<style>
.selectize-input {
  white-space: nowrap;
}
.selectize-dropdown {
  width: 450px !important;
}
#period + .shiny-input-select .selectize-dropdown {
  width: 150px !important;
}
</style>

```{r}
selectInput("period", "Period", names(periods))
```

```{r}
selectInput("research_area", "Research area", names(research_areas))
```



```{r}
textOutput("papers_evaluated")
```

::: {.callout-note}
**Note:** We now have 53 evaluated papers. The interactive visualizations are being updated and will be available shortly.
:::

## {.tabset}


### Research areas

#### Column

[Research areas](https://globalimpact.gitbook.io/the-unjournal-project-and-communication-space/policies-projects-evaluation-workflow/considering-projects/what-specific-areas-do-we-cover) of evaluated papers. 

```{r}
plotOutput("barplot_areas")
```


### Proposed papers

#### Column

Current status of all papers proposed for evaluation. We
don't currently record when papers were proposed, so this plot ignores the
time period chosen above.

```{r}
plotOutput("barplot_proposals")
```


### Overall Ratings

#### Column

Overall assessments for each paper, averaging evaluators' [research quality percentile scores](https://globalimpact.gitbook.io/the-unjournal-project-and-communication-space/policies-projects-evaluation-workflow/evaluation/guidelines-for-evaluators#overall-assessment).

```{r}
plotOutput("plot_ratings")
```

### Individual Criteria 

#### Column

Detailed evaluations of [different aspects of research quality](https://globalimpact.gitbook.io/the-unjournal-project-and-communication-space/policies-projects-evaluation-workflow/evaluation/guidelines-for-evaluators).


```{r}
plotOutput("plot_critera")
```


### Paper comparison

#### Column

```{r}
selectizeInput("comparison_papers", label = "Select papers to compare:", 
               choices = character(0), multiple = TRUE,
               options = list(
                 render = I("
{
    // display shortened names of selected papers
    item: function(item, escape) {
      const fullPaperName = item.value;
      const shortenedName = fullPaperName.length > 30 
        ? fullPaperName.substring(0, 30) + '...' 
        : fullPaperName;
      return '<div class=\"item\">' + escape(shortenedName) + '</div>';
    }
}
")
               ))

plotOutput("plot_comparison")
```



### Sources

#### Column

Papers can be chosen for evaluation by Unjournal staff,
submitted by authors or by a third party. 
Many come from our [internal project to evaluate NBER papers](https://globalimpact.gitbook.io/the-unjournal-project-and-communication-space/policies-projects-evaluation-workflow/considering-projects/direct-evaluation-track#the-case-for-this-direct-evaluation). Papers awaiting evaluation or selection are not included.

```{r}
plotOutput("barplot_sources")
```



```{r}
#| label: render-graphics
#| context: server


start_date <- reactive({
  periods[input$period]  
})

evals <- reactive({
  res_ratings |> 
    filter(
      row_created_date >= start_date(),
      input$research_area == "All" | main_cause_cat_abbrev == input$research_area
    )
})

evaled_research <- reactive({
  research |> 
    filter(status %in% "50_published evaluations (on PubPub, by Unjournal)") |> 
    # this applies the time filter
    semi_join(evals(), by = join_by(label_paper_title == research))
})

observe({
  titles <- evaled_research() |> pull(label_paper_title) |> sort()
  updateSelectizeInput(inputId = "comparison_papers",
                       choices = titles)
})

proposals <- reactive({
  research |> 
      filter(
      input$research_area == "All" | main_cause_cat_abbrev == input$research_area
    )
})

n_evals <- reactive(n_distinct(select(evals(), research, evaluator)))
n_papers <- reactive(nrow(evaled_research()))
all_cats <- reactive({
  evaled_research()$main_cause_cat_abbrev
})
n_areas <- reactive(n_distinct(all_cats()))

n_proposals <- reactive(nrow(proposals()))


output$papers_evaluated <- renderText({
  if (input$research_area == "All") {
    glue::glue("{n_evals()} evaluations of
    {n_papers()} papers in {n_areas()} distinct research areas")  
  } else {
    glue::glue("{n_evals()} evaluations of {n_papers()} papers") 
  }
})


output$barplot_areas <- renderPlot({
  data.frame(area = all_cats()) |>
    filter( ! is.na(area)) |>
    mutate(
      area = stringr::str_to_title(area),
      # Workaround to cope with long area categories. 
      # TODO: remove once fixed.
      # area = stringr::str_remove(area, "\\(.*\\)"),
      # area = stringr::str_remove(area, "\\[.*\\]"),
      area = forcats::fct_infreq(area),
      area = forcats::fct_rev(area),
    ) |>
    ggplot() +
      geom_bar(aes(y = area, fill = area)) +
      scale_x_continuous(
        breaks = scales::breaks_extended(Q = c(1, 5, 2, 4, 3))
      ) +
      labs(
        title = "Research areas of evaluated papers",
        x = NULL,
        y = NULL
      )
})


output$barplot_proposals <- renderPlot({
  proposals() |>
    mutate(
      Status = case_match(status,
        c("03_awaiting authors' consent (if needed) or update",
          "02_final consideration needed",
          "01_Needs prioritization/assessor rating")
          ~ "Under consideration",
        c("de-prioritized bc. of journal-publication status, authors' permission, age, etc.",
          "authors rejected/blocked us",
          "deprioritized -- low ratings/voting",
          "deprioritized -- out of scope or other subjective judgment or managerial concern",
          "deprioritized - no open access version",
          "Not a paper/project")
          ~ "Deprioritized",
        c("Applied and Policy Research Stream",
          "20_awaiting_evaluations",
          "contacting/awaiting_authors_response_to_evaluation",
          "In interim evaluation",
          "10_seeking_(more)_evaluators",
          "04_selected_choose_evaluation_manager")
          ~ "Accepted and under evaluation",
        c("50_published evaluations (on PubPub, by Unjournal)")
          ~ "Evaluation published",
        .default = "Unknown"
      ),
      Status = forcats::fct_relevel(Status, 
                                    "Under consideration", 
                                    "Deprioritized", 
                                    "Accepted and under evaluation", 
                                    "Evaluation published", "Unknown"),
      Status = forcats::fct_rev(Status)
      
    ) |>
    ggplot() +
      geom_bar(aes(y = Status, fill = Status)) +
      labs(
        title = "Status of papers proposed for evaluation",
        x = NULL,
        y = NULL
      )
})


output$plot_ratings <- renderPlot({
  evals() |>
    filter(criteria == "overall") |>
    mutate(
      area = stringr::str_to_title(main_cause_cat_abbrev),
      area = forcats::fct_infreq(area),
      area = forcats::fct_rev(area),
    ) |>
    ggplot(aes(y = area, x = middle_rating, color = area)) +
      geom_point() +
      stat_summary(fun = mean, geom = "point", na.rm = TRUE, shape = 3, size = 3) +
      scale_x_continuous(breaks = seq(0, 100, 20), limits = c(0, 100)) +
      theme(
        panel.grid.minor.x = element_blank(),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 14)
      ) +
      labs(
        title = "Overall ratings of papers by research area",
        subtitle = "Crosses show per-area averages",
        x = NULL,
        y = NULL
      )
})


criteria_plot_data <- reactive({
  evals() |>
    select(research, criteria, middle_rating) |>
    mutate(
      scale_0_to_5 = criteria %in% c("merits_journal", "journal_predict"),
      criteria = criteria_titles[criteria],
      criteria = forcats::fct_relevel(criteria, "Overall assessment"),
      criteria = forcats::fct_rev(criteria)
    ) |>
    filter(! is.na(criteria))
})

output$plot_critera <- renderPlot({

    ggp1 <- criteria_plot_data() |>
      filter(scale_0_to_5 == FALSE) |>
      ggplot(aes(y = criteria, x = middle_rating, color = criteria)) +
        geom_point() +
        stat_summary(fun = mean, geom = "point", na.rm = TRUE, shape = 3, size = 3) +
        scale_x_continuous(breaks = seq(0, 100, 20), limits = c(0, 100)) +
        theme(
          panel.grid.minor.x = element_blank(),
          axis.text.y = element_text(size = 10),
          axis.text.x = element_text(size = 10),
          plot.title = element_text(size = 14)
        ) +
        labs(
          title = "Answers to quantitative evaluation questions",
          subtitle = "Crosses show per-question averages",
          x = NULL,
          y = NULL
        )

  ggp2 <- criteria_plot_data() |>
    filter(scale_0_to_5 == TRUE) |>
    ggplot(aes(y = criteria, x = middle_rating, color = criteria)) +
    geom_point() +
    stat_summary(fun = mean, geom = "point", na.rm = TRUE, shape = 3, size = 3) +
    scale_x_continuous(breaks = seq(0, 5), limits = c(0, 5)) +
    theme(
      panel.grid.minor.x = element_blank(),
      axis.text.y = element_text(size = 10),
      axis.text.x = element_text(size = 10),
      plot.title = element_text(size = 14)
    ) +
    labs(
      x = NULL,
      y = NULL
    )

  cowplot::plot_grid(ggp1, ggp2, nrow = 2,
                     rel_heights = c(6, 3), align = "hv",
                     scale = 0.9)
})


output$plot_comparison <- renderPlot({
  evals() |>
    filter(
      research %in% input$comparison_papers
    ) |>
    summarize(.by = c(research, criteria),
              rating = mean(middle_rating, na.rm = TRUE)) |>
    mutate(
      rating = ifelse(criteria %in% c("merits_journal", "journal_predict"),
                      rating * 20, rating),
      criteria = criteria_titles[criteria],
      criteria = factor(criteria, levels = unique(criteria_titles)),
      criteria = forcats::fct_relevel(criteria, c("Overall assessment")),
      criteria = forcats::fct_rev(criteria),
      criteria = forcats::fct_relevel(criteria, c("Predicted journal tier",
                                                  "Deserved journal tier")),
      crit_grp = criteria %in% c("Predicted journal tier",
                                                  "Deserved journal tier")
    ) |>
    filter(! is.na(criteria)) |>
    ggplot(aes(x = rating, y = criteria, color = research,
               group = interaction(research, crit_grp))) +
      geom_point() +
      # hack to break the graph into two parts for other vs journal metrics:
      geom_hline(yintercept = 2.5, linewidth = 6, color = "white") +
      geom_line(orientation = "y") +
      # If we don't add this explicitly, the graph gets weird numbers on the
      # y axis instead of the criteria, when no papers are selected:
      scale_y_discrete(breaks = criteria_titles) +
      scale_x_continuous(limits = c(0, 100), position = "top",
                         breaks = seq(0, 100, 20),
                         sec.axis = sec_axis(~ . / 20, name = "Journal tier")) +
      labs(
        x = "Percentile rank",
        y = NULL
      ) +
      theme(
        panel.grid.major.y = element_line(color = "grey80"),
        panel.grid.minor.x = element_blank(),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 14),
        legend.position = "top",
        legend.direction = "vertical",
        legend.title = element_blank()
      )
})


output$barplot_sources <- renderPlot({
  evaled_research() |>
    mutate(
      Source = case_match(source_main,
        "internal-NBER"                ~ "Internal (NBER)",
        "internal-from-syllabus-agenda-policy-database" ~ "Internal",
        "suggested - internally"       ~ "Internal",
        "submitted (by author(s))"     ~ "Author",
        "suggested - externally"       ~ "Third party",
        "suggested - externally - NGO" ~ "Third party (NGO)",
        .default = "Unknown"
      ),
      Source = forcats::fct_infreq(Source),
      Source = forcats::fct_rev(Source)
    ) |>
    ggplot() +
      geom_bar(aes(y = Source, fill = Source)) +
      scale_x_continuous(
        breaks = scales::breaks_extended(Q = c(1, 5, 2, 4, 3))
      ) +
      labs(
        title = "Sources of evaluated papers",
        x = NULL,
        y = NULL
      )
})


```

