---
title: "Where Are Unjournal Papers Hosted? A Multi-Platform Analysis"
author: "David Reinstein with Claude Code assistance"
date: "2026-02-24"
categories: [data, infrastructure, open-science]
execute:
  echo: false
  warning: false
  message: false
freeze: auto
---

We analyzed where the 315 papers in The Unjournal's database are hosted across academic platforms and repositories. Understanding this distribution helps us identify opportunities for improving discoverability and integrating with scholarly infrastructure initiatives like [COAR Notify](https://notify.coar-repositories.org/).

## Key Findings

- **315 papers** analyzed (56 with published evaluations, 259 in our pipeline)
- **Average of 2.3 hosting locations** per paper — most papers appear on multiple platforms
- **Half (50%) of papers** appear on 2 or more platforms
- **26 papers (8%)** are on COAR Notify-enabled platforms (arXiv, OSF)
- **72 papers** appear across **47 institutional repositories**

## Multi-Platform Presence

A central finding of this analysis is that **papers typically exist on multiple platforms simultaneously**. This redundancy is valuable for discoverability but also means that COAR Notify integration could reach the same paper through multiple channels.

```{r}
#| label: multihost-setup
#| include: false

library(dplyr)
library(knitr)

multihost <- data.frame(
  Hosts = c("1 platform", "2 platforms", "3 platforms", "4 platforms", "5+ platforms"),
  Papers = c(157, 59, 33, 30, 35),
  Pct = c("50%", "19%", "10%", "10%", "11%")
)
```

```{r}
#| label: multihost-table

kable(multihost,
      col.names = c("Number of Hosting Locations", "Papers", "%"),
      caption = "Distribution of papers by number of hosting platforms")
```

### What Multi-Platform Hosting Looks Like

Papers with the most hosting locations (8+ platforms) tend to be high-impact research that gets deposited in institutional repositories, indexed by multiple services, and published in prominent journals. For example:

- **"A synthesis of evidence for policy from behavioural science during COVID-19"** (11 platforms): Nature, PubMed, LSE Research Online, Zurich Open Repository, eScholarship (UC system), and others
- **"Moral expansiveness around the world"** (11 platforms): SAGE journal, Edinburgh Research Explorer, Queen's University Belfast, HAL, and others
- **"Monetary incentives increase COVID-19 vaccinations"** (9 platforms): Science, PubMed, SERVAL (Lausanne), edoc (Basel), PubMed Central, and others

This pattern reflects how successful academic papers accumulate presence across: (1) the publisher's platform, (2) PubMed/PubMed Central for life sciences, (3) the authors' institutional repositories, and (4) discipline-specific indexes.

### Implications for COAR Integration

Multi-platform presence has strategic implications for **papers we have evaluated** (56 papers currently). COAR Notify would allow us to notify repositories when we publish an evaluation of a paper they host — this is relevant for evaluated papers, not papers still in our consideration pipeline.

1. **Notification efficiency**: An evaluated paper might be hosted on multiple COAR-enabled repositories, each of which could be notified about our evaluation
2. **Version tracking**: Different platforms may host different versions (submitted, accepted, published) — COAR Notify metadata could clarify which version was evaluated
3. **Reach optimization**: Focusing on high-coverage platforms (like arXiv or institutional repository networks) maximizes visibility for our evaluations

## Where Are Papers Hosted?

The table below shows how many papers appear on each platform, **counting all hosting locations** (not just primary URLs). Since half of our papers appear on 2+ platforms, the same paper may be counted under multiple rows.

```{r}
#| label: setup
#| include: false

# Data from our analysis (counting all hosting locations, papers may appear multiple times)
hosts <- data.frame(
  Platform = c("NBER", "PubMed (index)", "PubMed Central", "RePEc (index)",
               "SSRN", "arXiv", "Nature", "Institutional Repos", "OSF", "EconStor"),
  Papers = c(66, 58, 40, 39, 26, 18, 11, 72, 8, 6),
  Pct = c("21%", "18%", "13%", "12%", "8%", "6%", "3%", "23%", "3%", "2%"),
  Hosts_Papers = c("Yes", "No", "Yes", "No", "Yes", "Yes", "Yes", "Yes", "Yes", "Yes"),
  COAR_Status = c("Could implement", "N/A (index)", "Potential", "N/A (index)",
                  "Potential", "Enabled", "Via Crossref", "Via DSpace/EPrints", "Enabled", "In COAR directory")
)
```

```{r}
#| label: hosts-table

kable(hosts,
      col.names = c("Platform", "Papers", "%", "Hosts Papers?", "COAR Status"),
      caption = "Platform appearances across 315 papers (counting all hosting locations; papers may appear on multiple platforms)")
```

### An Important Distinction: Hosts vs. Indexes

Not all platforms that list papers actually *host* them:

- **Platforms that host papers** (store the PDFs): NBER, arXiv, SSRN, OSF, institutional repositories, PubMed Central
- **Indexes** (link to papers hosted elsewhere): RePEc, PubMed, DBLP

This distinction matters for infrastructure integration. [COAR Notify](https://coar-notify.net/), for example, can only integrate with platforms that actually host papers, since notifications need to link to the actual content.

## Repositories vs. Working Paper Series

A common question: what's the difference between a "repository" and a "working paper series"?

| Aspect | Repository | Working Paper Series |
|--------|-----------|---------------------|
| Examples | arXiv, Zenodo, OSF | NBER, CEPR, IZA |
| Who can deposit | Open to anyone (within scope) | Selective/affiliated authors |
| Hosts papers? | Yes | Yes |
| Could implement COAR? | Yes | Yes |

**Both host papers.** The difference is governance (open vs. selective), not technical capability. NBER, for instance, hosts PDFs and assigns DOIs just like arXiv does.

## COAR Notify: Current State and Opportunities

[COAR Notify](https://notify.coar-repositories.org/) is a protocol enabling repositories and review services to exchange notifications about peer review activities. This is directly relevant for The Unjournal as an overlay review service.

### Current COAR Readiness

```{r}
#| label: coar-table

coar_data <- data.frame(
  Status = c("Already COAR-enabled", "Potential targets", "Would need custom work", "Not applicable (indexes)"),
  Platforms = c("arXiv, OSF", "SSRN, PubMed Central, EconStor", "NBER, institutional repos", "RePEc, PubMed"),
  Papers = c("26 (8%)", "72 (23%)", "138 (44%)", "97 (31%)")
)

kable(coar_data,
      col.names = c("COAR Status", "Platforms", "Papers"),
      caption = "COAR Notify readiness of platforms hosting Unjournal papers")
```

### Strategic Pathways

Rather than reaching out to each institution individually, there are aggregation strategies:

1. **Platform-level adoption**: DSpace 8 now includes COAR Notify. If universities upgrade, dozens of institutional repositories gain COAR support simultaneously.

2. **Hub repositories**: [EconStor](https://www.econstor.eu/) (operated by ZBW, the German economics library) partners with 800+ institutions and is already in COAR's directory.

3. **Economics profession coordination**: Organizations like the AEA and NBER have coordinated infrastructure changes before.

## Institutional Repository Distribution

Our papers appear across 47 different university repositories:

**UK institutions** (25 papers): LSE, Cambridge, Oxford, Exeter, Warwick, Edinburgh, and others

**European institutions** (18 papers): Radboud, Copenhagen, Göttingen, Lund, Lausanne, and others

**US institutions** (15 papers): Chicago, Harvard, UC system (eScholarship), Minnesota, and others

**Asia-Pacific institutions** (8 papers): Monash, NUS, Hong Kong universities

Many of these use DSpace or EPrints, platforms that support or are developing COAR Notify integration.

## Comparison: Evaluated Papers vs. Full Pipeline

This comparison matters because **only evaluated papers can currently benefit from COAR Notify** — we would notify repositories about our published evaluations of these papers.

```{r}
#| label: comparison-table

comparison <- data.frame(
  Platform = c("NBER", "SSRN", "arXiv", "OSF", "PubMed Central", "Institutional repos"),
  Evaluated = c("17 (30%)", "3 (5%)", "4 (7%)", "2 (4%)", "2 (4%)", "~10"),
  All = c("66 (21%)", "26 (8%)", "18 (6%)", "8 (3%)", "40 (13%)", "72")
)

kable(comparison,
      col.names = c("Platform", "56 Evaluated Papers", "315 All Papers"),
      caption = "Platform appearances: evaluated papers vs. full pipeline (counting all hosting locations)")
```

**Key observation for COAR**: Among our 56 evaluated papers, approximately 6 (11%) are on COAR-enabled platforms (4 on arXiv, 2 on OSF). These papers could immediately benefit from COAR Notify integration.

The broader pipeline shows more diversity in hosting platforms, with notably more papers in PubMed Central and SSRN than among the currently evaluated set.

## Data Limitations

Our analysis used OpenAlex, Semantic Scholar, and CrossRef APIs. Some limitations:

- **RePEc under-counted**: OpenAlex links only ~25% of NBER papers to their RePEc entries (actual coverage is nearly 100%)
- **Newer papers**: May not yet be fully indexed
- **Personal websites**: Cannot be automatically discovered

The actual multi-platform presence is likely 20-30% higher than reported.

## Implications for The Unjournal

**For our 56 evaluated papers:**

1. **Immediate integration**: Papers we've evaluated that are hosted on arXiv or OSF (COAR-enabled platforms) could benefit from COAR Notify now — repositories could be notified of our published evaluations

2. **Multi-platform notifications**: Since evaluated papers often appear on multiple platforms, a single evaluation could potentially trigger notifications to several repositories

**For future evaluations:**

3. **Author guidance**: Encouraging authors to deposit on arXiv, OSF, or Zenodo (in addition to NBER/SSRN) would expand COAR coverage for future evaluations

4. **Strategic partnerships**: EconStor/ZBW is a natural partner for economics overlay review — they're already in the COAR directory

5. **Platform advocacy**: Supporting DSpace/EPrints COAR adoption helps the broader ecosystem and increases coverage for institutional repositories

## Methodology

We queried OpenAlex, Semantic Scholar, and CrossRef APIs for each paper, using title matching (threshold >0.80) and DOI lookup where available. Results were cross-referenced and deduplicated. The analysis code and raw data are available in our [data repository](https://github.com/unjournal/unjournaldata).

---

*This analysis was conducted in February 2026 to support discussions with [COAR](https://www.coar-repositories.org/) about infrastructure integration for overlay peer review.*
