---
title: "Where Are Unjournal Papers Hosted? A Multi-Platform Analysis"
author: "David Reinstein with Claude Code assistance"
date: "2026-02-24"
categories: [data, infrastructure, open-science]
execute:
  echo: false
  warning: false
  message: false
freeze: auto
---

We analyzed where the 315 papers in The Unjournal's database are hosted across academic platforms and repositories. Understanding this distribution helps us identify opportunities for improving discoverability and integrating with scholarly infrastructure initiatives like [COAR Notify](https://notify.coar-repositories.org/).

## Key Findings

**For our 57 evaluated papers** (where COAR Notify could link evaluations to repositories):

- **22 papers (39%)** are on SSRN — a major potential integration point if SSRN implements COAR Notify
- **17 papers (30%)** are on NBER, which could implement COAR Notify
- **7 papers (12%)** are on COAR-enabled platforms (arXiv, OSF) — ready for integration now
- **15 papers (26%)** are hosted only on organization websites or published journals — no clear COAR path
- **Average 2.3 hosting locations** per paper — many appear on multiple platforms

**For the full database (315 papers including 258 in pipeline):**

- **Half (50%) of papers** appear on 2 or more platforms
- **19 papers verified** in institutional repositories via OpenAlex (3 evaluated, 16 pipeline) — DSpace is the most common platform (11 papers)
- **49 papers** are on SSRN (many cross-posted from NBER or other sources)
- More diversity in platforms: notably more in PubMed Central and IZA

*Note on SSRN:* Many papers are cross-posted to SSRN even when their primary URL points elsewhere (e.g., NBER, personal sites). Our analysis found 22 evaluated papers and 27 pipeline papers on SSRN — far more than the 16 with SSRN as their primary link.

## Multi-Platform Presence

A central finding of this analysis is that **papers typically exist on multiple platforms simultaneously**. This redundancy is valuable for discoverability but also means that COAR Notify integration could reach the same paper through multiple channels.

```{r}
#| label: multihost-setup
#| include: false

library(dplyr)
library(knitr)

multihost <- data.frame(
  Hosts = c("1 platform", "2 platforms", "3 platforms", "4 platforms", "5+ platforms"),
  Papers = c(157, 59, 33, 30, 35),
  Pct = c("50%", "19%", "10%", "10%", "11%")
)
```

```{r}
#| label: multihost-table

kable(multihost,
      col.names = c("Number of Hosting Locations", "Papers", "%"),
      caption = "Distribution of papers by number of hosting platforms")
```

### What Multi-Platform Hosting Looks Like

Papers with the most hosting locations (8+ platforms) tend to be high-impact research that gets deposited in institutional repositories, indexed by multiple services, and published in prominent journals. For example:

- **"A synthesis of evidence for policy from behavioural science during COVID-19"** (11 platforms): Nature, PubMed, LSE Research Online, Zurich Open Repository, eScholarship (UC system), and others
- **"Moral expansiveness around the world"** (11 platforms): SAGE journal, Edinburgh Research Explorer, Queen's University Belfast, HAL, and others
- **"Monetary incentives increase COVID-19 vaccinations"** (9 platforms): Science, PubMed, SERVAL (Lausanne), edoc (Basel), PubMed Central, and others

This pattern reflects how successful academic papers accumulate presence across: (1) the publisher's platform, (2) PubMed/PubMed Central for life sciences, (3) the authors' institutional repositories, and (4) discipline-specific indexes.

### Implications for COAR Integration

Multi-platform presence has strategic implications for **papers we have evaluated** (57 papers currently). COAR Notify would allow us to notify repositories when we publish an evaluation of a paper they host — this is relevant for evaluated papers, not papers still in our consideration pipeline.

1. **Notification efficiency**: An evaluated paper might be hosted on multiple COAR-enabled repositories, each of which could be notified about our evaluation
2. **Version tracking**: Different platforms may host different versions (submitted, accepted, published) — COAR Notify metadata could clarify which version was evaluated
3. **Reach optimization**: Focusing on high-coverage platforms (like arXiv or institutional repository networks) maximizes visibility for our evaluations

## Where Are Papers Hosted?

The table below shows how many papers appear on each platform, **counting all hosting locations** (not just primary URLs). Since half of our papers appear on 2+ platforms, the same paper may be counted under multiple rows.

```{r}
#| label: setup
#| include: false

# Data from our analysis (counting all hosting locations, papers may appear multiple times)
# Includes both detected counts and estimated actuals (accounting for API under-reporting)
# SSRN counts updated Feb 2026 after deep-dive cross-referencing analysis
# Ordered by All_Detected (prevalence)
hosts <- data.frame(
  Platform = c("Other*", "Institutional Repos", "NBER", "PubMed (index)", "SSRN",
               "PubMed Central", "RePEc (index)", "arXiv", "Nature portfolio",
               "OSF", "EconStor", "IZA", "CEPR"),
  Eval_Detected = c(15, 3, 17, 4, 22, 2, 14, 4, 1, 3, 0, 0, 0),
  Eval_Est = c("15", "3‡", "17", "4", "22+", "2", "25-30†", "4", "1", "3", "0-1", "0", "0"),
  All_Detected = c(113, 19, 66, 58, 49, 40, 39, 18, 18, 14, 5, 5, 1),
  All_Est = c("113", "19+‡", "66", "58", "50-60", "40", "80-100†", "18", "18", "14", "6-8", "5", "1"),
  Hosts_Papers = c("Various", "Yes", "Yes", "No", "Yes", "Yes", "No", "Yes", "Yes", "Yes", "Yes", "Yes", "Yes"),
  COAR_Status = c("—", "Via DSpace/EPrints", "Could implement", "N/A (index)", "Potential",
                  "Potential", "N/A (index)", "Enabled", "Via Crossref", "Enabled",
                  "In COAR directory", "Could implement", "Could implement")
)
```

```{r}
#| label: hosts-table

kable(hosts,
      col.names = c("Platform", "57 Evaluated (detected)", "57 Evaluated (est.)",
                    "315 All (detected)", "315 All (est.)", "Hosts Papers?", "COAR Status"),
      caption = "Platform appearances (counting all hosting locations; papers may appear on multiple platforms)")
```

**Notes on the table:**

- *SSRN cross-posting:* Many NBER papers are cross-posted to SSRN. Of our 49 total papers on SSRN, only 16 have SSRN as primary URL — the rest are cross-posted from NBER or other sources. Examples include "Cash Transfers for Child Development" ([SSRN 4709146](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=4709146)), "AI and Economic Growth" ([SSRN 3053718](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3053718)), and pipeline papers like "Longer-run Economic Consequences of Pandemics" ([SSRN 3594164](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=3594164)). Full mapping available in `data/ssrn_crosspost_mapping.csv`.
- *‡Institutional repos (verified via OpenAlex):* We comprehensively queried OpenAlex for all 57 evaluated papers and 150 pipeline papers using both DOI and title search. Found 3 evaluated and 16 pipeline papers in institutional repositories. Primary platforms: DSpace (11), UK repos (4), EU repos (4), eScholarship (3), HAL (2), PURE (1).
- *†RePEc estimates:* OpenAlex links only ~25% of NBER papers to their RePEc entries; actual RePEc coverage of NBER papers is nearly 100%. Our detected counts significantly underestimate RePEc presence.
- **Other*** includes: organization websites (e.g., Rethink Priorities, ALLFED), personal hosting (Squarespace, Dropbox), published OA journals not in major indexes, and ResearchGate. These 15 evaluated papers and 113 pipeline papers are hosted somewhere but not on the major platforms listed above.

### An Important Distinction: Hosts vs. Indexes

Not all platforms that list papers actually *host* them:

- **Platforms that host papers** (store the PDFs): NBER, arXiv, SSRN, OSF, institutional repositories, PubMed Central
- **Indexes** (link to papers hosted elsewhere): RePEc, PubMed, DBLP

This distinction matters for infrastructure integration. [COAR Notify](https://coar-notify.net/), for example, can only integrate with platforms that actually host papers, since notifications need to link to the actual content.

## Repositories vs. Working Paper Series

A common question: what's the difference between a "repository" and a "working paper series"?

| Aspect | Repository | Working Paper Series |
|--------|-----------|---------------------|
| Examples | arXiv, Zenodo, OSF | NBER, CEPR, IZA |
| Who can deposit | Open to anyone (within scope) | Selective/affiliated authors |
| Hosts papers? | Yes | Yes |
| Could implement COAR? | Yes | Yes |

**Both host papers.** The difference is governance (open vs. selective), not technical capability. NBER, for instance, hosts PDFs and assigns DOIs just like arXiv does.

## COAR Notify: Current State and Opportunities

[COAR Notify](https://notify.coar-repositories.org/) is a protocol enabling repositories and review services to exchange notifications about peer review activities. This is directly relevant for The Unjournal as an overlay review service.

### Current COAR Readiness

```{r}
#| label: coar-table

coar_data <- data.frame(
  Status = c("Already COAR-enabled", "Potential targets (SSRN, PMC, EconStor)", "Would need custom work", "Not applicable (indexes)", "Other hosting"),
  Platforms = c("arXiv, OSF", "SSRN (49 papers), PubMed Central, EconStor", "NBER, IZA, CEPR, institutional repos", "RePEc, PubMed", "Org websites, journals, personal"),
  Evaluated = c("7 (12%)", "24 (42%)", "24 (42%)", "18 (32%)*", "15 (26%)"),
  All = c("32 (10%)", "91 (29%)", "144 (46%)", "97 (31%)*", "113 (36%)")
)

kable(coar_data,
      col.names = c("COAR Status", "Platforms", "57 Evaluated", "315 All Papers"),
      caption = "COAR Notify readiness of platforms hosting Unjournal papers")
```

*Note: Papers may appear in multiple categories since many are on 2+ platforms. The "indexes" category (RePEc, PubMed) counts papers that appear there but doesn't reflect COAR potential since indexes don't host papers.*

### Strategic Pathways

Rather than reaching out to each institution individually, there are aggregation strategies:

1. **Platform-level adoption**: DSpace 8 now includes COAR Notify. If universities upgrade, dozens of institutional repositories gain COAR support simultaneously.

2. **Hub repositories**: [EconStor](https://www.econstor.eu/) (operated by ZBW, the German economics library) partners with 800+ institutions and is already in COAR's directory.

3. **Economics profession coordination**: Organizations like the AEA and NBER have coordinated infrastructure changes before.

## Institutional Repository Distribution

"Institutional repositories" encompasses both the **software platforms** universities use and the **specific university repositories** themselves. Understanding both is important for COAR Notify integration strategy.

### Repository Software Platforms

Universities run institutional repositories using various software platforms. Platform-level COAR adoption would enable many repositories simultaneously:

```{r}
#| label: software-platforms

software_data <- data.frame(
  Platform = c("DSpace", "EPrints", "PURE (Elsevier)", "Digital Commons", "Figshare", "HAL"),
  Type = c("Open-source", "Open-source", "Commercial", "Commercial", "Commercial", "National (France)"),
  Usage = c("2000+ institutions globally (MIT, Cambridge, Edinburgh...)",
            "500+ institutions, strong in UK (LSE, Southampton, White Rose consortium...)",
            "Nordic/Dutch universities (Copenhagen, Lund, Radboud...)",
            "US universities, especially liberal arts colleges",
            "Newer entrant (Monash, some UK universities)",
            "All French research institutions (single integration = national coverage)"),
  COAR_Status = c("DSpace 8 includes COAR Notify!",
                  "Community developing support",
                  "Would need Elsevier buy-in",
                  "Would need vendor buy-in",
                  "Would need vendor buy-in",
                  "Potential - national coordination")
)

kable(software_data,
      col.names = c("Platform", "Type", "Typical Users", "COAR Status"),
      caption = "Repository software platforms and their COAR Notify potential")
```

**Verified institutional repo presence (via OpenAlex):** We comprehensively queried OpenAlex for all 57 evaluated papers and 150 pipeline papers (using both DOI lookup and title search). Results:

- **3 evaluated papers** found in institutional repositories
- **16 pipeline papers** found in institutional repositories
- **19 papers total** with verified institutional repo presence

**By platform:**

| Platform | Evaluated | Pipeline | Total |
|----------|-----------|----------|-------|
| DSpace | 3 | 8 | 11 |
| UK inst. repos (Warwick, Edinburgh, etc.) | 0 | 4 | 4 |
| EU inst. repos (Radboud, Lund, etc.) | 0 | 4 | 4 |
| eScholarship (UC) | 1 | 2 | 3 |
| HAL (France) | 0 | 2 | 2 |
| PURE | 0 | 1 | 1 |
| US inst. repos | 0 | 1 | 1 |

*Note:* Some papers appear in multiple repositories. For example, "Monetary incentives increase COVID-19 vaccinations" was found in DSpace, UChicago Knowledge, Lund LUP, and Lausanne SERVAL simultaneously.

### Specific University Repositories

Based on our OpenAlex verification, papers appear in these specific institutional repositories:

**UK institutions:** LSE Research Online, Cambridge Apollo, Oxford ORA, Warwick WRAP, Edinburgh Research Explorer, White Rose (Leeds/Sheffield/York consortium), UCL Discovery, Exeter ORE

**European institutions:** Radboud Repository, Copenhagen CURIS, Göttingen, Lund LUP, Lausanne SERVAL, Basel edoc, HAL (all French institutions), Zurich Open Repository

**US institutions:** eScholarship (UC system — 10 campuses with single integration), Harvard DASH, UChicago Knowledge, Minnesota Conservancy, Stanford SearchWorks

**Asia-Pacific institutions:** Monash, NUS ScholarBank, HKU Hub, University of Melbourne

### Strategic Implications

1. **Platform-level wins multiply reach**: DSpace 8 adoption spreading means dozens of repositories gain COAR support simultaneously
2. **Consortium repositories**: White Rose (3 UK universities), eScholarship (10 UC campuses), HAL (all French research) — single integrations with multiplicative effect
3. **Commercial platforms need vendor buy-in**: PURE and Digital Commons require Elsevier and bepress cooperation respectively

## Key Differences: Evaluated Papers vs. Full Pipeline

The main hosting table above shows both groups side-by-side. Key differences:

**COAR-ready evaluated papers**: Among our 57 evaluated papers, approximately 7 (12%) are on COAR-enabled platforms (4 on arXiv, 3 on OSF). These papers could immediately benefit from COAR Notify integration.

**Pipeline shows more diversity**: The broader pipeline has notably more papers in PubMed Central (40 vs. 2), reflecting a broader range of research fields beyond economics. Institutional repository presence is similar across both groups when accounting for cross-posting.

**Working paper series**: NBER dominates our evaluated papers (30%), while the pipeline includes more representation from IZA (5 papers) and one CEPR paper. These economics working paper series could potentially implement COAR Notify.

**"Other" hosting**: 15 evaluated papers (26%) and 113 pipeline papers (36%) are hosted only on organization websites, personal hosting, or published journals not in major indexes. This represents a significant portion that wouldn't benefit from platform-level COAR integration.

## Data Limitations

Our analysis used OpenAlex, Semantic Scholar, and CrossRef APIs. Some limitations:

- **RePEc under-counted**: OpenAlex links only ~25% of NBER papers to their RePEc entries (actual coverage is nearly 100%)
- **Newer papers**: May not yet be fully indexed
- **Personal websites**: Cannot be automatically discovered

The actual multi-platform presence is likely 20-30% higher than reported.

## Implications for The Unjournal

**For our 57 evaluated papers:**

1. **SSRN is a major opportunity**: 22 evaluated papers (39%) are on SSRN. If SSRN implements COAR Notify, this single integration would cover more of our papers than any other platform except NBER.

2. **Immediate integration**: 7 papers on arXiv or OSF (COAR-enabled platforms) could benefit from COAR Notify now — repositories could be notified of our published evaluations

3. **Multi-platform notifications**: Since evaluated papers often appear on multiple platforms, a single evaluation could potentially trigger notifications to several repositories

**For future evaluations:**

4. **Author guidance**: Encouraging authors to deposit on arXiv, OSF, or Zenodo (in addition to NBER/SSRN) would expand COAR coverage for future evaluations

5. **Strategic partnerships**: EconStor/ZBW is a natural partner for economics overlay review — they're already in the COAR directory

6. **Platform advocacy**: Supporting DSpace/EPrints COAR adoption helps the broader ecosystem and increases coverage for institutional repositories

## Methodology

We queried OpenAlex, Semantic Scholar, and CrossRef APIs for each paper, using title matching (threshold >0.80) and DOI lookup where available. Results were cross-referenced and deduplicated. The SSRN cross-posting analysis used targeted web searches to verify availability on SSRN for papers hosted elsewhere. The analysis code and raw data are available in our [data repository](https://github.com/unjournal/unjournaldata).

---

*This analysis was conducted in February 2026 to support discussions with [COAR](https://www.coar-repositories.org/) about infrastructure integration for overlay peer review.*
